---
title: "Group Assignment: Replicating Ouss and Stevenson (2023)"
author: "Group Members: Julian Schmidt, Laurens Mauz, Isa Afridi, Lukas Simon"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_depth: 3
    number_sections: false
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# Load libraries here (e.g., library(tidyverse), library(fixest), library(DoubleML))
```

## 1.1 Differences-in-Differences
Test markdown
Test function

```{r}
covars_to_test <- c("defendantageatarrest", "male", "defendantisblack", 
                    "prior_FTA", "prior_9y", "past_fel", "felony")
```
did_cov_data <- CashBail %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = as.numeric(prior_fel > 0),
    Released = if_else(Jail3days == 0, 1, 0)
  )
```

## 1.2 Custom Double ML DiD Function
## 1.3 Replication of Ouss and Stevenson (2023)
### 1.3.1 Overview
### 1.3.2 Descriptives
#### Cash Bail System
#### Policy Reform
#### Target Parameters and Controls
### 1.3.3 Replication Results
#### DiD TWFE
#### DiD Double ML
#### Heterogeneity
### 1.3.4 Reflection on DiD Assumptions
#### Policy Context
#### Event Study DiD
#### Endogenous Responses
#### Channels of Cash-Bail Reform Effects on Violations
### 1.3.5 Feedback