---
title: 'Group Assignment: Replicating Ouss and Stevenson (2023)'
author: 'Group Members: Julian Schmidt, Laurens Mauz, Isa Afridi, Lukas Simon'
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    number_sections: false
    theme: cosmo
    highlight: tango
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(fixest)
library(haven)
library(gt)
library(ggh4x)
library(modelsummary)
library(gtsummary)
library(DoubleML)
library(kableExtra)
library(cowplot)
library(NuisanceParameters)
library(hdm)
library(broom)
```

```{r}
master_data <- read_stata("CashBail.dta")
master_data$baildate <- as.Date(master_data$baildate, origin = "1960-01-01")
```

## 1.1 Differences-in-Differences

## 1.2 Custom Double ML DiD Function

## 1.3 Replication of Ouss and Stevenson (2023)

### 1.3.1 Overview

### 1.3.2 Descriptives

#### Descriptive Statistics Table

```{r}
desc_stats_set <- master_data %>%
  filter(Post == 0) %>% 
  mutate(
    defendantzipmedianhouseholdincom = round(defendantzipmedianhouseholdincom, 0),
    PctBelPov = PctBelPov/100,
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = if_else(coalesce(prior_fel, 0) > 0, 1, 0),
    group_label = if_else(`EligibleOffense` == 1, "Eligible cases", "Ineligible cases")
  ) %>%
  select(
    group_label,
    defendantageatarrest, male, defendantisblack, Hisp, White, 
    defendantzipmedianhouseholdincom, PctBelPov, PD, felony, 
    prior_FTA, prior_9y, past_fel, 
    ROR, Nonmonetary, Unsecured, Bail_5000, Bail_HIGH, Denied, 
    Jail3days, hasfta, Rec6mo
  )

####
desc_stats_final <- desc_stats_set %>%
  tbl_summary(
    by = group_label,
    # forcing R to treat 0/1 variables as continuous so it can calculate mean 
    type = list(everything() ~ "continuous"), 
    label = list(
      defendantageatarrest ~ "Age",
      male ~ "Male",
      defendantisblack ~ "Black",
      Hisp ~ "Hispanic",
      White ~ "White",
      defendantzipmedianhouseholdincom ~ "Median household income in zip code",
      PctBelPov ~ "Fraction below poverty in zip code",
      PD ~ "Public Defender",
      felony ~ "Felony",
      prior_FTA ~ "Has a prior FTA",
      prior_9y ~ "Has a past conviction",
      past_fel ~ "Has a past felony conviction",
      ROR ~ "ROR",
      Nonmonetary ~ "Supervised release",
      Unsecured ~ "Unsecured monetary",
      Bail_5000 ~ "Secured bail up to 5,000",
      Bail_HIGH ~ "Secured bail over 5,000",
      Denied ~ "Denied bail",
      Jail3days ~ "Jail (3+ nights)",
      hasfta ~ "FTA",
      Rec6mo ~ "Recidivism"
    ),
    statistic = all_continuous() ~ "{mean}",
    digits = all_continuous() ~ 2,
    missing = "no"
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_spanning_header(all_stat_cols() ~ "**Pre-Reform Cases**") %>%
  bold_labels()

desc_stats_final %>%
  as_gt() %>%
  tab_row_group(label = "Misconduct", rows = 20:21) %>%
  tab_row_group(label = "Pre-trial Conditions", rows = 13:19) %>%
  tab_row_group(label = "Defendant Characteristics", rows = 1:12)
```

#### Cash Bail System

```{r}
data_trends_ov <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),
    
    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

biweekly_avg <- data_trends_ov %>%
  group_by(W2, EligibleOffense) %>%
  summarise(
    ROR = mean(ROR, na.rm = TRUE),
    Jail = mean(Jail3days, na.rm = TRUE),
    FTA = mean(hasfta, na.rm = TRUE),
    Recidivism = mean(Rec6mo, na.rm = TRUE),
    .groups = "drop"
  ) %>%
 
  filter(W2 >= -12 & W2 <= 12 & W2 != 0) %>%
  pivot_longer(cols = c(ROR, Jail, FTA, Recidivism), names_to = "Outcome", values_to = "Rate")

scales_y <- list(
  Outcome == "ROR" ~ scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)),
  Outcome == "Jail" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "FTA" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "Recidivism" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1))
)

biweekly_avg <- biweekly_avg %>%
  mutate(Outcome = factor(Outcome, levels = c("ROR", "Jail", "FTA", "Recidivism")))

# ggplot(biweekly_avg, aes(x = W2, y = Rate)) +
#   geom_point(color = "grey70", size = 1.5) +
#   
#   
#   geom_smooth(data = subset(biweekly_avg, W2 < 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_smooth(data = subset(biweekly_avg, W2 > 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
#   
#   facet_wrap(~Outcome, scales = "free_y") +
#   
#   facetted_pos_scales(y = scales_y) +
#   
#   scale_x_continuous(breaks = seq(-12, 12, 4), 
#                      labels = function(x) x / 2) +
#   
#   theme_minimal() +
#   labs(x = "Months relative to Feb. 22",
#        y = "Bi-weekly Rate") +
#   theme(panel.grid.minor = element_blank())
        
```

```{r}
# Figure 1a: ROR Time Trend
p_eligible <- ggplot(subset(biweekly_avg, Outcome == "ROR"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  
  theme_minimal() +
  labs(title = "ROR",
       subtitle = "(Eligible Cases)",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly ROR Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )
```

```{r}

# Figure 1c: ROR for Ineligible Cases
data_inel <- master_data %>%
  filter(EligibleOffense == 0) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),

    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

# Bi-weekly ROR Rate
biweekly_ror_inel <- data_inel %>%
  group_by(W2) %>%
  summarise(Rate = mean(ROR, na.rm = TRUE), .groups = "drop") %>%
  filter(W2 >= -12 & W2 <= 11)

p_ineligible <- ggplot(biweekly_ror_inel, aes(x = W2, y = Rate)) +
 
  geom_point(color = "grey70", size = 1.2) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 < 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 >= 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "ROR", 
       subtitle = "(Ineligible Cases)",
       x = "Months relative to Feb. 22", 
       y = "Bi-weekly ROR rate") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )
  
```

## Figure 1a: ROR Time Trend & Figure 1b: ROR for Ineligible Cases

```{r}

plot_grid(p_eligible, p_ineligible,
          ncol = 2, nrow = 1)

```

#### Policy Reform

#### Target Parameters and Controls

### 1.3.3 Replication Results

#### DiD TWFE

```{r}
## Data Preparation
did_twfe <- master_data %>%
  mutate(
    yq_numeric = as.numeric(as.factor(yq)),
    # Creating past_fel from prior_fel 
    past_fel = as.numeric(prior_fel > 0),
    
    # Creating Denied 
    Denied = as.numeric(initialbailtype == "Denied"),
    
    # Addressing NAs
    income_missing = as.numeric(is.na(defendantzipmedianhouseholdincom)),
    poverty_missing = as.numeric(is.na(PctBelPov)),
    
    defendantzipmedianhouseholdincom = ifelse(is.na(defendantzipmedianhouseholdincom), 
                                              mean(defendantzipmedianhouseholdincom, na.rm = TRUE), 
                                              defendantzipmedianhouseholdincom),
    
    PctBelPov = ifelse(is.na(PctBelPov), 
                       mean(PctBelPov, na.rm = TRUE), 
                       PctBelPov),
    
    # Re-calculate or ensure PostElig is correct
    PostElig = EligibleOffense * Post,
  
    off = as.factor(off),
    yq = as.factor(yq),
    DOW = as.factor(DOW), 
    CommissionerName_input = as.factor(CommissionerName_input) 
  )

covariates <- c("defendantageatarrest", "male", "defendantisblack", "Hisp", 
                "White", "defendantzipmedianhouseholdincom", "PctBelPov", 
                "PD", "felony", "prior_FTA", "prior_9y", "past_fel", "income_missing", "poverty_missing")

# Column 1: ROR (No controls)
res_ror_1 <- feols(ROR ~ PostElig + EligibleOffense + Post, 
                   cluster = ~off, data = did_twfe)

# Column 2: ROR (With controls and Fixed Effects)
res_ror_2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, 
                   cluster = ~off, data = did_twfe)

# Column 3: Jail (No controls)
res_jail_3 <- feols(Jail3days ~ PostElig + EligibleOffense + Post, 
                    cluster = ~off, data = did_twfe)

# Column 4: Jail (With controls and Fixed Effects)
res_jail_4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, 
                    cluster = ~off, data = did_twfe)

## Pre-period Means & Models
mean_ror_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(ROR, na.rm = TRUE)) %>% pull(mean)

mean_jail_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Jail3days, na.rm = TRUE)) %>% pull(mean)

# Model (1): ROR No Controls
m1 <- feols(ROR ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): ROR With Controls & Fixed Effects
m2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

# Model (3): Jail No Controls
m3 <- feols(Jail3days ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Jail With Controls & Fixed Effects
m4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

## SUMMARY TABLE

table_3 <- etable(m1, m2, m3, m4,
       headers = list("ROR" = 2, "Jail" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_ror_pre, 3), round(mean_ror_pre, 3), 
                             round(mean_jail_pre, 3), round(mean_jail_pre, 3))
       ),
      
       digits = 4, 
       digits.stats = "r0", 
       fitstat = ~n,     
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

mean_fta_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(hasfta, na.rm = TRUE)) %>% pull(mean)

mean_rec_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Rec6mo, na.rm = TRUE)) %>% pull(mean)

# Model (1): FTA No Controls
m5_1 <- feols(hasfta ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): FTA With Controls & Fixed Effects
m5_2 <- feols(hasfta ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

# Model (3): Recidivism No Controls
m5_3 <- feols(Rec6mo ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Recidivism With Controls & Fixed Effects
m5_4 <- feols(Rec6mo ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

#  SUMMARY TABLE 5
table_5 <- etable(m5_1, m5_2, m5_3, m5_4,
       headers = list("FTA" = 2, "Recidivism" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_fta_pre, 3), round(mean_fta_pre, 3), 
                             round(mean_rec_pre, 3), round(mean_rec_pre, 3))
       ),
       digits = 4, 
       digits.stats = "r0",
       fitstat = ~n,
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

```

## Table 3: DiD Estimates of policy effect on ROR and Jail & Table 5: DiD Estimates of policy effect on FTA and Recidivism

```{r}
table_6 <- etable(
  m2, m4, m5_2, m5_4,
  dict = c(PostElig = "Eligible x Post 02/21"),
  keep = "Eligible x Post 02/21",
  extralines = list(
    "Mean dep. var." = c(round(mean_ror_pre, 3),
                         round(mean_jail_pre, 3),
                         round(mean_fta_pre, 3),
                         round(mean_rec_pre, 3))
  ),
  digits = 4,
  digits.stats = "r0",
  fitstat = ~n,
  style.df = style.df(depvar.title = "", fixef.suffix = ""),
  notes = "Standard errors, clustered at the offense level, are in parentheses."
)

# Render with kableExtra for a clean academic look
library(kableExtra)

table_6 %>%
  as.data.frame() %>%
  kbl(caption = "Table x: DiD Estimates of ROR, Jail, FTA, and Recidivism",
      booktabs = TRUE, 
      align = "lcccc",
      # This removes the extra column name row kable sometimes adds
      col.names = c("Characteristic", "ROR", "Jail", "FTA", "Recidivism")) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```

#### DiD Double ML

We observe $(Y_i, D_i, T_i, X_i)$ where $D_i \in \{0,1\}$ is the treatment variable and $T_i \in \{0,1\}$ is the time variable. The parameter of interest is the ATT (average treatment effect on the treated): $$\theta_0 = E[(Y(1) - Y(0)| D=1 ]$$ identified in Cross Sectional DiD under parallel trend and standard overlap assumption.

We thereby use a doubly robust AIPW scores to get the parameter estimate $\hat{\theta}$ as well as the influence function $\hat{\psi}_i(\theta)$. In order to be able to use AIPW we need to estimate the nuisance parameter propensity $$m_0(X) = P(D=1|X) \quad \forall D\in \{0,1\}$$ as well as the outcome regressions for each cell $(d,t)$ which is $$g_{dt,0}(X) = E[Y|D=d, T=t,X]$$ by the given nuisance parameter function. Since the given function does not allow to conditioning on $(D,T,X)$ we need constructed a workaround such that we can get $\hat{g}_{11}(X),\hat{g}_{01}(X), \hat{g}_{10}(X), \hat{g}_{00}(X)$. We estimate that by turning $(D,T)$ into a 4-category label `DT_combined <- 2 * D + Tt`.

We then constructed $\psi_a$ and $\psi_b$ based on the score provided by the [DoubleML](https://docs.doubleml.org/stable/guide/scores.html#id4) documentation. Using this we solve proceeded by solving the linear moment equation: $$ 
E[\psi_a \theta + \psi_b] = 0 \Rightarrow \hat{\theta} = - \frac{E_i[\psi_b]}{E_i[\psi_a]}
$$ Afterward we constructed the influence function $$ IF_i = \frac{-1}{E[\psi_a]} \cdot \psi_i(\hat{\theta})$$ and used it to get clustered standard errors.

#### Heterogeneity

Let $D_i \in \{0,1\}$ denote treatment status, $T_i \in \{0,1\}$ a time indicator, and $G_i \in \mathcal G$ a discrete group variable. We are interested in the group-specific average treatment effect on the treated $$
\tau_g \;=\; \mathbb E\!\left[ Y_i(1) - Y_i(0) \mid D_i=1,\, G_i=g \right].
$$ We start from a doubly robust score for the overall ATT, which yields $$
\theta^{ATT}
=
-\frac{\mathbb E[\psi_b]}{\mathbb E[\psi_a]}.
$$

To obtain a group-specific ATT, we do not condition on $D_i=1$. Instead, we define a group-weighted moment condition using the indicator $I_{g,i} = \mathbf 1\{G_i=g\}$: $$
\mathbb E\!\left[ I_{g,i}\,(\psi_{a,i}\,\tau_g + \psi_{b,i}) \right] = 0.
$$ Solving for $\tau_g$ gives $$
\boxed{
\tau_g
=
-\frac{\mathbb E[I_g\,\psi_b]}{\mathbb E[I_g\,\psi_a]}
}
$$ which is the GATT estimator implied by the same score that identifies the ATT.

#### DML DiD estimates of ATT and GATT

```{r}

doubleML_did_cs_gate <- function(data, y_var, d_var, t_var, x_vars, methods, cf = 5,
                                 feature_eng = FALSE, bin_cut = 0.02, corr_cut = 0.9,
                                 cluster_var = NULL,
                                 heterogeneity = FALSE, gate_var = NULL) {
  
  Y  <- data[[y_var]]
  D  <- as.numeric(data[[d_var]])
  Tt <- as.numeric(data[[t_var]])
  X_0 <- as.matrix(data[, x_vars])
  n  <- length(Y)
  
  if (!is.null(cluster_var)) {
    C <- data[[cluster_var]]
  }
  
  if (feature_eng) {
    X_expanded <- design_matrix(X_0, int = "all", int_d = 2,
                                poly = "all", poly_d = 2)
    X <- data_screen(X_expanded, treat = D,
                     bin_cut = bin_cut, corr_cut = corr_cut,
                     quiet = TRUE)
  } else {
    X <- X_0
  }
  
  # Kombiniertes Treatment: (D,T)
  DT_combined <- 2 * D + Tt
  
  # ---- 1) m(X)
  np_m <- nuisance_parameters(
    NuPa = "D.hat", X = X, D = D,
    methods = methods, cf = cf,
    stacking = "short", stratify = TRUE, ensemble_type = "nnls",
    quiet = TRUE
  )
  m_hat <- as.numeric(np_m$nuisance_parameters$D.hat)
  
  # ---- 2) g(d,t,X)
  np_g <- nuisance_parameters(
    NuPa = "Y.hat.d",
    X = X,
    Y = Y,
    D = DT_combined,
    methods = methods,
    cf = cf,
    stacking = "short",
    stratify = TRUE,
    ensemble_type = "nnls",
    quiet = TRUE
  )
  
  g00_hat <- np_g$nuisance_parameters$Y.hat.d[, 1]  # D=0,T=0
  g01_hat <- np_g$nuisance_parameters$Y.hat.d[, 2]  # D=0,T=1
  g10_hat <- np_g$nuisance_parameters$Y.hat.d[, 3]  # D=1,T=0
  g11_hat <- np_g$nuisance_parameters$Y.hat.d[, 4]  # D=1,T=1
  
  # ---- 3) Score (ATT)
  psi_a <- -D / mean(D)
  
  psi_b <- D / mean(D) * (g11_hat - g10_hat - (g01_hat - g00_hat)) +
    (Tt * D / mean(D * Tt)) * (Y - g11_hat) -
    (D * (1 - Tt) / mean(D * (1 - Tt))) * (Y - g10_hat) -
    (m_hat * (1 - D) * Tt / (1 - m_hat)) /
    mean(m_hat * (1 - D) * Tt / (1 - m_hat)) * (Y - g01_hat) +
    (m_hat * (1 - D) * (1 - Tt) / (1 - m_hat)) /
    mean(m_hat * (1 - D) * (1 - Tt) / (1 - m_hat)) * (Y - g00_hat)
  
  theta_hat <- -mean(psi_b) / mean(psi_a)
  
  psi <- psi_a * theta_hat + psi_b
  influence_fun <- -1 / mean(psi_a) * psi
  
  if (is.null(cluster_var)) {
    se_hat <- sqrt(mean(psi^2) / (mean(psi_a)^2 * n))
    clustered <- FALSE
  } else {
    clusters <- unique(C)
    cluster_sums_squared <- rep(NA_real_, length(clusters))
    index <- 1
    for (cl in clusters) {
      cluster_sums_squared[index] <- sum(influence_fun[C == cl])^2
      index <- index + 1
    }
    var_hat <- 1 / n * sum(cluster_sums_squared)
    se_hat <- sqrt(var_hat / n)
    clustered <- TRUE
  }
  
  # ---- 4) Return object (base)
  out <- list(
    coefficients = c(ATT = theta_hat),
    se = se_hat,
    clustered = clustered,
    t = theta_hat / se_hat,
    p = 2 * (1 - pnorm(abs(theta_hat / se_hat))),
    ci_lower = theta_hat - 1.96 * se_hat,
    ci_upper = theta_hat + 1.96 * se_hat,
    N = n,
    N_treated = sum(D),
    N_control = sum(1 - D),
    N_post = sum(Tt),
    N_pre = sum(1 - Tt),
    y_var = y_var,
    d_var = d_var,
    t_var = t_var,
    x_vars = x_vars,
    n_covariates = ncol(X),
    cf = cf,
    methods = names(methods),
    feature_eng = feature_eng,
    cluster_var = cluster_var
  )
  
  # ------ 5) GATE  -------
  
  if (heterogeneity) {
    
    if (is.null(gate_var)) stop("heterogeneity = TRUE requires gate_var")
    
    G <- as.factor(data[[gate_var]])
    groups <- levels(G)
    
    # containers
    tau_hat <- rep(NA_real_, length(groups))
    names(tau_hat) <- groups
    
    IF_mat <- matrix(NA_real_, nrow = n, ncol = length(groups))
    colnames(IF_mat) <- groups
    
    for (j in seq_along(groups)) {
      g <- groups[j]
      I_g <- as.numeric(G == g)
      
      denom <- mean(I_g * psi_a)
      if (abs(denom) < 1e-12) {
        stop(paste0("Group ", g, " has near-zero denom E[I_g*psi_a]."))
      }
      
      num <- mean(I_g * psi_b)
      tau_hat[g] <- - num / denom
      
      # Influence function for tau_g
      psi_g <- I_g * (psi_a * tau_hat[g] + psi_b)
      IF_mat[, j] <- -1 / denom * psi_g
    }
    
    # SEs using the same logic as ATT, applied per group
    gate_se <- rep(NA_real_, length(groups))
    names(gate_se) <- groups
    
    if (is.null(cluster_var)) {
      for (j in seq_along(groups)) {
        gate_se[j] <- sqrt(mean(IF_mat[, j]^2) / n)
      }
      gate_clustered <- FALSE
    } else {
      clusters <- unique(C)
      for (j in seq_along(groups)) {
        infl <- IF_mat[, j]
        cluster_sums_squared <- rep(NA_real_, length(clusters))
        k <- 1
        for (cl in clusters) {
          cluster_sums_squared[k] <- sum(infl[C == cl])^2
          k <- k + 1
        }
        var_hat_g <- 1 / n * sum(cluster_sums_squared)
        gate_se[j] <- sqrt(var_hat_g / n)
      }
      gate_clustered <- TRUE
    }
    
    gate_t <- tau_hat / gate_se
    gate_p <- 2 * (1 - pnorm(abs(gate_t)))
    
    out$gate <- list(
      tau = tau_hat,
      se = gate_se,
      clustered = gate_clustered,
      t = gate_t,
      p = gate_p,
      ci_lower = tau_hat - 1.96 * gate_se,
      ci_upper = tau_hat + 1.96 * gate_se,
      groups = groups,
      N = n,
      gate_var = gate_var,
      influence_function = IF_mat
    )
  }
  class(out) <- "doubleML_did_cs_gate"
  return(out)
}


```


```{r}
# data preparation (was don alread in did_twfe, but I addedyq_numeric)
CashBail <- read_stata("CashBail.dta") %>%
  mutate(
    yq_numeric = as.numeric(as.factor(yq)),
    # Creating past_fel from prior_fel
    past_fel = as.numeric(prior_fel > 0),
    
    # Creating Denied
    Denied = as.numeric(initialbailtype == "Denied"),
    
    # Addressing NAs
    income_missing = as.numeric(is.na(defendantzipmedianhouseholdincom)),
    poverty_missing = as.numeric(is.na(PctBelPov)),
    
    defendantzipmedianhouseholdincom = ifelse(
      is.na(defendantzipmedianhouseholdincom),
      mean(defendantzipmedianhouseholdincom, na.rm = TRUE),
      defendantzipmedianhouseholdincom
    ),
    
    PctBelPov = ifelse(is.na(PctBelPov), mean(PctBelPov, na.rm = TRUE), PctBelPov),
    
    PostElig = EligibleOffense * Post,
    
    off = as.factor(off),
    yq = as.factor(yq),
    DOW = as.factor(DOW),
    CommissionerName_input = as.factor(CommissionerName_input),
    # add a group_variable that indicates whether a defendant is white and not hispanic,         black and not hispanic or hispanic
    ethnicity = factor(
      if_else(
        defendantisblack == 1 & Hisp == 0,
        "black_not_hisp",
        if_else(White == 1 & Hisp == 0, "white_not_hisp", 
                if_else(Hisp == 1, "hisp", NA))
      ),
      levels = c("black_not_hisp", "white_not_hisp", "hisp")
    )
  )
CashBail <- filter(CashBail, !is.na(ethnicity)) 


```

```{r}
# defining inputs
outcomes_table3 <- c("ROR", "Jail3days")
outcomes_table5 <- c("hasfta", "Rec6mo")
all_outcomes <- c(outcomes_table3, outcomes_table5)

covariates <- c(
  "defendantisblack",
  "Hisp",
  "White",
  "defendantageatarrest",
  "male",
  "defendantzipmedianhouseholdincom",
  "PctBelPov",
  "PD",
  "felony",
  "prior_FTA",
  "prior_9y",
  "past_fel",
  "income_missing",
  "poverty_missing"
)

# ML Methods
my_methods <- list(D.hat = list(
  m1 = list(method = "ranger_prop", tune = "fold"),
  m2 = list(method = "logit")
),

Y.hat.d = list(
  m1 = list(method = "ranger", tune = "fold"),
  m2 = list(method = "lasso")
))
```


```{r eval=FALSE}
# Store the estimation results for each outcome variable
dml_gate_results <- list()
set.seed(1234) 

# (ATT is automatically computed as part of GATT estimation)
for(y in all_outcomes) {
  message("Running ATT estimation for outcome: ", y)
  
  dml_results[[y]] <- doubleML_did_cs_gate(
    data = CashBail,
    y_var = y,
    d_var = "EligibleOffense",
    t_var = "Post",
    x_vars = covariates,
    methods = my_methods,
    cf = 5,              
    cluster_var = "off",
    heterogeneity = TRUE,
    gate_var = "ethnicity",
    feature_eng = TRUE
  )
}
```

#### Table ?: ATT Results 
```{r eval=FALSE}

# Extract ATT from first GATT variable results (they all have the same ATT)
final_dml_table <- do.call(rbind, lapply(all_outcomes, function(y) {
  res <- dml_gate_results[[y]] # Take ATT from first gate_var
  data.frame(
    Outcome = y,
    ATT = res$coefficients["ATT"],
    Std_Error = res$se,
    t_stat = res$t,
    p_val = res$p,
    Lower_CI = res$ci_lower,
    Upper_CI = res$ci_upper
  )
}))

# make the  ATT table more beautiful
final_dm_rtable_beautiful <- final_dml_table %>%
  select(!c(Std_Error, t_stat)) %>%
  gt() %>%
  tab_header(
    title = "DiD ATT Results"
  ) %>%
  cols_label(
    Outcome = "Outcome",
    ATT = "ATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Outcome, ATT, p_val, Lower_CI, Upper_CI), decimals = 3) 

final_dm_rtable_beautiful
```


#### GATT Results Summary
```{r eval=FALSE}
# Summary Table for GATT, dependent variable = hasfta
groups <- c("black_not_hisp", "white_not_hisp", "hisp")
summary_table_hasfta <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$hasfta$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$hasfta$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$hasfta$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$hasfta$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$hasfta$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$hasfta$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_hasfta) <- NULL


summary_table_hasfta_formatted <- summary_table_hasfta %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for hasfta"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 



```



```{r eval=FALSE}
# Summary Table for GATT, dependent variable = Rec6mo
summary_table_Rec6mo <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_Rec6mo) <- NULL


summary_table_Rec6mo_formatted <- summary_table_Rec6mo %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for Rec6mo"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 

```


```{r eval=FALSE}
# Summary tabble DiD GATT, dep. variable = ROR
summary_table_ROR <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$ROR$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$ROR$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$ROR$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$ROR$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$ROR$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$ROR$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_ROR) <- NULL


summary_table_ROR_formatted <- summary_table_ROR %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for ROR"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 




```


```{r eval=FALSE}
# Summary Table for DiiD GATT, dep. variable = Jail3days
summary_table_Jail3days <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$Jail3days$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$Jail3days$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$Jail3days$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$Jail3days$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$Jail3days$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$Jail3days$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_Jail3days) <- NULL


summary_table_Jail3days_formatted <- summary_table_Jail3days %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for Jail3days"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 



```

# Table ?: GATT for Dependent Variablehasfta
```{r eval=FALSE}
summary_table_hasfta_formatted 
```


# Table ?: GATT for Dependent Variable Rec6mo
```{r eval=FALSE}
summary_table_Rec6mo_formatted 
```


# Table ?: GATT for Dependent Variable ROR
```{r eval=FALSE}
summary_table_ROR_formatted 
```


# Table ?: GATT for Dependent Variable Jail
```{r eval=FALSE}
summary_table_Jail3days_formatted 
```


### 1.3.4 Reflection on DiD Assumptions

#### Policy Context

#### Event Study DiD

#### Endogenous Responses

#### Channels of Cash-Bail Reform Effects on Violations

## Figure 2: Jail Time Trend

```{r}
ggplot(subset(biweekly_avg, Outcome == "Jail"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "Jail",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly Jail Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

### Table x: Characterisitcs Composition

```{r}
covars_to_test <- c("defendantageatarrest", "male", "defendantisblack",
                    "prior_FTA", "prior_9y", "past_fel", "felony")
did_cov_data <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = as.numeric(prior_fel > 0),
    Released = if_else(Jail3days == 0, 1, 0)
  )

did_results <- map_df(covars_to_test, function(var) {
  form <- as.formula(paste(var, "~ Post * Released"))
  lm(form, data = did_cov_data) %>%
    tidy() %>%
    filter(term == "Post:Released") %>%
    mutate(Variable = var)
})

did_cov_table <- did_results %>%
  select(Variable, estimate, std.error, p.value) %>%
  mutate(
    Status = if_else(p.value > 0.05, "Stable", "Shift Detected")
  )


did_cov_table_formatted <- did_cov_table %>%
  gt() %>%
  tab_header(
    title = "DiD Test for Compositional Shifts"
  ) %>%
  cols_label(
    Variable = "Characteristic",
    estimate = "Interaction (Post × Released)",
    std.error = "Std. Error",
    p.value = "P-Value"
  ) %>%
  fmt_number(columns = c(estimate, std.error, p.value), decimals = 3) %>%
  tab_style(
    style = cell_text(color = "darkgreen", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "Stable")
  )

did_cov_table_formatted
```

### 1.3.5 Feedback
