---
title: 'Group Assignment: Replicating Ouss and Stevenson (2023)'
author: 'Group Members: Julian Schmidt, Laurens Mauz, Isa Afridi, Lukas Simon'
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    number_sections: false
    theme: cosmo
    highlight: tango
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(fixest)
library(haven)
library(gt)
library(ggh4x)
library(modelsummary)
library(gtsummary)
library(DoubleML)
library(kableExtra)
library(cowplot)
library(NuisanceParameters)
library(hdm)
library(broom)
```

```{r}
master_data <- read_stata("CashBail.dta")
master_data$baildate <- as.Date(master_data$baildate, origin = "1960-01-01")
```

## 1.1 Differences-in-Differences
Difference-in-Differences is a causal inference method used on observed data (i.e. not experimental setting) to estimate causal effects. It does so by comparing the changes in outcomes over time between a population that is enrolled in a program (the Treatment Group) and a population that is not (the Control Group). The intuition is the following: We want to know what would have happened to the treated units had they not been treated. Since we cannot observe this counterfactual, we use the trend of the control group as a proxy for what the treated group's trend would have been.

### Identification

$$
\begin{aligned}
\mathbb{E}[Y_2 \mid W = 1] - \mathbb{E}[Y_1 \mid W = 1] &\stackrel{\text{consistency}}{=} \mathbb{E}[Y_2(1) \mid W = 1] - \mathbb{E}[Y_1(1) \mid W = 1] \\
&= \mathbb{E}[Y_2(1) - Y_1(1) \mid W = 1] + \color{blue}{\mathbb{E}[Y_2(0) - Y_2(0) \mid W = 1]} + \color{blue}{\mathbb{E}[Y_1(0) - Y_1(0) \mid W = 1]} \\
&= \underbrace{\mathbb{E}[Y_2(1) - Y_2(0) \mid W = 1]}_{\tau_{ATT}} + \underbrace{\mathbb{E}[Y_2(0) - Y_1(0) \mid W = 1]}_{\text{counterfactual time trend}} - \underbrace{\mathbb{E}[Y_1(1) - Y_1(0) \mid W = 1]}_{\text{anticipation effect}}
\end{aligned}
$$

Unlike standard Cross-Sectional OLS, which relies on the unconfoundedness assumption ($Y(0), Y(1) \perp W \mid X$), DiD relaxes this requirement. It allows for selection bias—where individuals choose treatment based on unobserved characteristics—provided that these unobserved factors are time-invariant, and the parallel trends and no-anticipation effect assumptions hold. 

The former requires trends in the absence of treatment to be parallel, i.e. $\mathbb{E}[Y_2(0) - Y_1(0) \mid W=1]=\mathbb{E}[Y_2(0) - Y_1(0) \mid W=0]$. 

The no-anticipation effect assumption requires that the outcomes of the treatment group are only affected by the treatment after the treatment is given, i.e., $$\mathbb{E}[Y_1(1) - Y_1(0) \mid W=1] = 0$$.

The downside of not needing the unconfoundedness assumption is that we can only identify the ATT, not the ATU or ATE. This is the case because we do not assume parallel trends under treatment, i.e., $\mathbb{E}[Y_2(1) - Y_1(1) \mid W=1]=\mathbb{E}[Y_2(1) - Y_1(1) \mid W=0]$. In other words, the control group might be different in some unobserved confounders which make treatment effects heterogeneous from the treatment group.

In practice researchers often rely on conditional parallel trends. This assumption acknowledges that treatment and control groups may have different compositions (e.g., age or gender distributions) that naturally lead to diverging trends. Therefore trends are parallel only after adjusting for these covariates. Furthermore, with repeated cross-sectional data researcher control for relevant covariates in order to account for potential group-composition shifts between periods.

## 1.2 Custom Double ML DiD Function

## 1.3a Replication of Ouss and Stevenson (2023)

### 1.3a.2 Descriptives

#### Cash Bail System
The paper “Does Cash Bail Deter Misconduct?” (Ouss and Stevenson, 2023) investigates how cash bails affect the likelihood of a defendant violating the bail conditions. Cash bailout refers to a financial collateral the defendant must pay to remain out of jail while they await their trial. If the defendant fulfils the bail conditions they get their collateral back. The conditions which need to be fulfilled in order for the defendant to get the cash bail back after the court trials are over:

- **Appearing to all hearings**
- **Not committing another crime**

If an individual violates either of these conditions, they lose the bail collateral. Furthermore, if an individual does not even pay the collateral in the beginning, they are detained in jail until the court trials are over.

The cash bail system thus is supposed to ensure individuals appear for the hearings and don’t commit further crimes through a monetary incentive. Thus individuals who do not pay the bail collateral are jailed, because the lack of monetary incentive would make it more likely for them to violate the conditions.

#### Policy Reform
The reform analysed in this paper is a change in the policy of Philadelphia’s public prosecutors, who in February 2018 recommended to no longer require cash-bail for many low-level offenses. The motivation of this and similar reforms in other jurisdictions is to protect economically vulnerable, often minority groups which cannot afford to post cash-bail. If the cash-bail system is successful in motivating individuals to not violate the bail conditions, then releasing defendants without financial incentive should lead to more violations. This paper thus uses DiD to investigate whether eligible defenders (i.e. defenders eligible for the new bail policy, who committed low-level crimes) who now no longer need to post bail are after the reform more likely to miss court hearings and commit further crimes.

Importantly, the reform entailed the Philadelphia prosecutor only recommending this change in bail policy instead of simply deciding, because the decision whether and how much cash-bail is required is made by bail magistrates, who have the ultimate discretion here. While this discretion with magistrates makes it thus imaginable that this reform did not actually change anything, figure 1 shows that the share of defendants released without cash-bail (the correct legal term being Release on own Recognizance, or ROR) increased significantly, by around 15-20 percentage points. In the replication section, we will test this increase in ROR also through a DiD design.


```{r}
data_trends_ov <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),
    
    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

biweekly_avg <- data_trends_ov %>%
  group_by(W2, EligibleOffense) %>%
  summarise(
    ROR = mean(ROR, na.rm = TRUE),
    Jail = mean(Jail3days, na.rm = TRUE),
    FTA = mean(hasfta, na.rm = TRUE),
    Recidivism = mean(Rec6mo, na.rm = TRUE),
    .groups = "drop"
  ) %>%
 
  filter(W2 >= -12 & W2 <= 12 & W2 != 0) %>%
  pivot_longer(cols = c(ROR, Jail, FTA, Recidivism), names_to = "Outcome", values_to = "Rate")

scales_y <- list(
  Outcome == "ROR" ~ scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)),
  Outcome == "Jail" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "FTA" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "Recidivism" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1))
)

biweekly_avg <- biweekly_avg %>%
  mutate(Outcome = factor(Outcome, levels = c("ROR", "Jail", "FTA", "Recidivism")))

# ggplot(biweekly_avg, aes(x = W2, y = Rate)) +
#   geom_point(color = "grey70", size = 1.5) +
#   
#   
#   geom_smooth(data = subset(biweekly_avg, W2 < 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_smooth(data = subset(biweekly_avg, W2 > 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
#   
#   facet_wrap(~Outcome, scales = "free_y") +
#   
#   facetted_pos_scales(y = scales_y) +
#   
#   scale_x_continuous(breaks = seq(-12, 12, 4), 
#                      labels = function(x) x / 2) +
#   
#   theme_minimal() +
#   labs(x = "Months relative to Feb. 22",
#        y = "Bi-weekly Rate") +
#   theme(panel.grid.minor = element_blank())
        
```


```{r}
# Figure 1a: ROR Time Trend
p_eligible <- ggplot(subset(biweekly_avg, Outcome == "ROR"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  
  theme_minimal() +
  labs(title = "ROR",
       subtitle = "(Eligible Cases)",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly ROR Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) + 
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )
```


```{r}

# Figure 1c: ROR for Ineligible Cases
data_inel <- master_data %>%
  filter(EligibleOffense == 0) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),

    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

# Bi-weekly ROR Rate
biweekly_ror_inel <- data_inel %>%
  group_by(W2) %>%
  summarise(Rate = mean(ROR, na.rm = TRUE), .groups = "drop") %>%
  filter(W2 >= -12 & W2 <= 11)

p_ineligible <- ggplot(biweekly_ror_inel, aes(x = W2, y = Rate)) +
 
  geom_point(color = "grey70", size = 1.2) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 < 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 >= 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "ROR", 
       subtitle = "(Ineligible Cases)",
       x = "Months relative to Feb. 22", 
       y = "Bi-weekly ROR rate") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )
  
```

## Figure 1a: ROR Time Trend & Figure 1b: ROR for Ineligible Cases

```{r}

plot_grid(p_eligible, p_ineligible,
          ncol = 2, nrow = 1)

```
The authors argue that the DA led reform thus did not affect bail policy through a change in formal rules, but rather through a change in norms.



#### Target Parameters and Controls
The authors scraped the data for this analysis off the Philadelphia court websites. The auxiliary target parameters are:

- **Release on Recognizance (ROR):** Whether a defendant is released without monetary bail or supervisory conditions. As explained in the prior section,  measuring ROR is necessary to confirm that the policy actually changed bail-setting behavior.
- **Pretrial Detention:** Defined as spending at least three nights in jail immediately following the initial bail hearing. The DiD will later show that the amount of individuals detained before trial was not affected by the reform. This will be used to argue that any reform affects on the FTA and recidivism variables must be due to the monetary incentive channel and not the detainment channel.

The target parameters of the main DiD analysis are:

- **Failure to Appear (FTA):** Defined as equal to one if the defendant fails to appear for at least one court date associated with the case.
- **Recidivism:** Defined as equal to one if the defendant is charged with a new offense within six months of the bail hearing.

Both of these main target parameters make up the conditions which cannot be violated if the defendant wants to keep their bail money.

To account for case and defendant characteristics, the authors include the following controls in the DiD analyses:

- **Defendant Demographics:** Race, gender, and age at arrest.
- **Criminal History:** The presence and number of prior FTAs and prior convictions (limited to a nine-year window before the bail hearing),,.
- **Case Characteristics:** The type of offense (aggregated to the 23 most common offenses plus a catchall category), the grade of offense, and whether the defendant was represented by a public defender.
- **Procedural Factors:** The specific bail magistrate overseeing the hearing, the day of the week, and the magistrate work shift.
- **Standard errors:** in these analyses are clustered at the offense level to account for correlations within offense types.


### 1.3a.3 Replication Results

#### DiD TWFE
$$
(1) \quad Y_{i} = \alpha + \beta Post_{i} + \delta Post_{i} \times Eligible_{i} + \lambda Eligible_{i} + \theta \mathbf{X}_{i} + \epsilon_{i}.
$$

Equation (1) specifies the paper’s DiD regression, where Y is one of the 4 target parameters, i indicates the specific case, Post the post-treatment period and Eligible that an individual is part of the treatment group affected by the reform. The covariates used are the described above in the section target parameters and controls. The main coefficient of interest is δ. The authors do not include individual level fixed effects because we are in a cross-sectional setting. Instead we have one time fixed effect (Post) and one group fixed effect (eligible).
We replicate the TWFE in the paper ourselves to double check that we have the data set up correctly in order for the double ML estimation to be comparable.
```{r}
## Data Preparation
did_twfe <- master_data %>%
  mutate(
    yq_numeric = as.numeric(as.factor(yq)),
    # Creating past_fel from prior_fel 
    past_fel = as.numeric(prior_fel > 0),
    
    # Creating Denied 
    Denied = as.numeric(initialbailtype == "Denied"),
    
    # Addressing NAs
    income_missing = as.numeric(is.na(defendantzipmedianhouseholdincom)),
    poverty_missing = as.numeric(is.na(PctBelPov)),
    
    defendantzipmedianhouseholdincom = ifelse(is.na(defendantzipmedianhouseholdincom), 
                                              mean(defendantzipmedianhouseholdincom, na.rm = TRUE), 
                                              defendantzipmedianhouseholdincom),
    
    PctBelPov = ifelse(is.na(PctBelPov), 
                       mean(PctBelPov, na.rm = TRUE), 
                       PctBelPov),
    
    # Re-calculate or ensure PostElig is correct
    PostElig = EligibleOffense * Post,
  
    off = as.factor(off),
    yq = as.factor(yq),
    DOW = as.factor(DOW), 
    CommissionerName_input = as.factor(CommissionerName_input) 
  )

covariates <- c("defendantageatarrest", "male", "defendantisblack", "Hisp", 
                "White", "defendantzipmedianhouseholdincom", "PctBelPov", 
                "PD", "felony", "prior_FTA", "prior_9y", "past_fel", "income_missing", "poverty_missing")

# Column 1: ROR (No controls)
res_ror_1 <- feols(ROR ~ PostElig + EligibleOffense + Post, 
                   cluster = ~off, data = did_twfe)

# Column 2: ROR (With controls and Fixed Effects)
res_ror_2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, 
                   cluster = ~off, data = did_twfe)

# Column 3: Jail (No controls)
res_jail_3 <- feols(Jail3days ~ PostElig + EligibleOffense + Post, 
                    cluster = ~off, data = did_twfe)

# Column 4: Jail (With controls and Fixed Effects)
res_jail_4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, 
                    cluster = ~off, data = did_twfe)

## Pre-period Means & Models
mean_ror_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(ROR, na.rm = TRUE)) %>% pull(mean)

mean_jail_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Jail3days, na.rm = TRUE)) %>% pull(mean)

# Model (1): ROR No Controls
m1 <- feols(ROR ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): ROR With Controls & Fixed Effects
m2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

# Model (3): Jail No Controls
m3 <- feols(Jail3days ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Jail With Controls & Fixed Effects
m4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

## SUMMARY TABLE

table_3 <- etable(m1, m2, m3, m4,
       headers = list("ROR" = 2, "Jail" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_ror_pre, 3), round(mean_ror_pre, 3), 
                             round(mean_jail_pre, 3), round(mean_jail_pre, 3))
       ),
      
       digits = 4, 
       digits.stats = "r0", 
       fitstat = ~n,     
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

mean_fta_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(hasfta, na.rm = TRUE)) %>% pull(mean)

mean_rec_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Rec6mo, na.rm = TRUE)) %>% pull(mean)

# Model (1): FTA No Controls
m5_1 <- feols(hasfta ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): FTA With Controls & Fixed Effects
m5_2 <- feols(hasfta ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

# Model (3): Recidivism No Controls
m5_3 <- feols(Rec6mo ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Recidivism With Controls & Fixed Effects
m5_4 <- feols(Rec6mo ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

#  SUMMARY TABLE 5
table_5 <- etable(m5_1, m5_2, m5_3, m5_4,
       headers = list("FTA" = 2, "Recidivism" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_fta_pre, 3), round(mean_fta_pre, 3), 
                             round(mean_rec_pre, 3), round(mean_rec_pre, 3))
       ),
       digits = 4, 
       digits.stats = "r0",
       fitstat = ~n,
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

```

#### Table 3: DiD Estimates of policy effect on ROR and Jail & Table 5: DiD Estimates of policy effect on FTA and Recidivism

```{r}
table_6 <- etable(
  m2, m4, m5_2, m5_4,
  dict = c(PostElig = "Eligible x Post 02/21"),
  keep = "Eligible x Post 02/21",
  extralines = list(
    "Mean dep. var." = c(round(mean_ror_pre, 3),
                         round(mean_jail_pre, 3),
                         round(mean_fta_pre, 3),
                         round(mean_rec_pre, 3))
  ),
  digits = 4,
  digits.stats = "r0",
  fitstat = ~n,
  style.df = style.df(depvar.title = "", fixef.suffix = ""),
  notes = "Standard errors, clustered at the offense level, are in parentheses."
)

# Render with kableExtra for a clean academic look
library(kableExtra)

table_6 %>%
  as.data.frame() %>%
  kbl(caption = "Table x: DiD Estimates of ROR, Jail, FTA, and Recidivism",
      booktabs = TRUE, 
      align = "lcccc",
      # This removes the extra column name row kable sometimes adds
      col.names = c("Characteristic", "ROR", "Jail", "FTA", "Recidivism")) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```
We will not interpret these TWFE DiD results by themselves here, but rather use them to benchmark our Double ML result in the following section.

#### DiD Double ML
For the Double ML estimation we use the same dataset with the same target parameters and covariates as in the TWFE implementation.   
Here now a brief explanation of how we constructed the custom Double ML function:
We observe $(Y_i, D_i, T_i, X_i)$ where $D_i \in \{0,1\}$ is the treatment variable and $T_i \in \{0,1\}$ is the time variable. The parameter of interest is the ATT (average treatment effect on the treated): $$\theta_0 = E[(Y(1) - Y(0)| D=1 ]$$ identified in Cross Sectional DiD under parallel trend and standard overlap assumption.

We thereby use a doubly robust AIPW scores to get the parameter estimate $\hat{\theta}$ as well as the influence function $\hat{\psi}_i(\theta)$. In order to be able to use AIPW we need to estimate the nuisance parameter propensity $$m_0(X) = P(D=1|X) \quad \forall D\in \{0,1\}$$ as well as the outcome regressions for each cell $(d,t)$ which is $$g_{dt,0}(X) = E[Y|D=d, T=t,X]$$ by the given nuisance parameter function. Since the given function does not allow to conditioning on $(D,T,X)$ we need constructed a workaround such that we can get $\hat{g}_{11}(X),\hat{g}_{01}(X), \hat{g}_{10}(X), \hat{g}_{00}(X)$. We estimate that by turning $(D,T)$ into a 4-category label `DT_combined <- 2 * D + Tt`.

We then constructed $\psi_a$ and $\psi_b$ based on the score provided by the [DoubleML](https://docs.doubleml.org/stable/guide/scores.html#id4) documentation. Using this we solve proceeded by solving the linear moment equation: $$ 
E[\psi_a \theta + \psi_b] = 0 \Rightarrow \hat{\theta} = - \frac{E_i[\psi_b]}{E_i[\psi_a]}
$$ Afterward we constructed the influence function $$ IF_i = \frac{-1}{E[\psi_a]} \cdot \psi_i(\hat{\theta})$$ and used it to get clustered standard errors.


```{r}

doubleML_did_cs_gate <- function(data, y_var, d_var, t_var, x_vars, methods, cf = 5,
                                 feature_eng = FALSE, bin_cut = 0.02, corr_cut = 0.9,
                                 cluster_var = NULL,
                                 heterogeneity = FALSE, gate_var = NULL) {
  
  Y  <- data[[y_var]]
  D  <- as.numeric(data[[d_var]])
  Tt <- as.numeric(data[[t_var]])
  X_0 <- as.matrix(data[, x_vars])
  n  <- length(Y)
  
  if (!is.null(cluster_var)) {
    C <- data[[cluster_var]]
  }
  
  if (feature_eng) {
    X_expanded <- design_matrix(X_0, int = "all", int_d = 2,
                                poly = "all", poly_d = 2)
    X <- data_screen(X_expanded, treat = D,
                     bin_cut = bin_cut, corr_cut = corr_cut,
                     quiet = TRUE)
  } else {
    X <- X_0
  }
  
  # Kombiniertes Treatment: (D,T)
  DT_combined <- 2 * D + Tt
  
  # ---- 1) m(X)
  np_m <- nuisance_parameters(
    NuPa = "D.hat", X = X, D = D,
    methods = methods, cf = cf,
    stacking = "short", stratify = TRUE, ensemble_type = "nnls",
    quiet = TRUE
  )
  m_hat <- as.numeric(np_m$nuisance_parameters$D.hat)
  
  # ---- 2) g(d,t,X)
  np_g <- nuisance_parameters(
    NuPa = "Y.hat.d",
    X = X,
    Y = Y,
    D = DT_combined,
    methods = methods,
    cf = cf,
    stacking = "short",
    stratify = TRUE,
    ensemble_type = "nnls",
    quiet = TRUE
  )
  
  g00_hat <- np_g$nuisance_parameters$Y.hat.d[, 1]  # D=0,T=0
  g01_hat <- np_g$nuisance_parameters$Y.hat.d[, 2]  # D=0,T=1
  g10_hat <- np_g$nuisance_parameters$Y.hat.d[, 3]  # D=1,T=0
  g11_hat <- np_g$nuisance_parameters$Y.hat.d[, 4]  # D=1,T=1
  
  # ---- 3) Score (ATT)
  psi_a <- -D / mean(D)
  
  psi_b <- D / mean(D) * (g11_hat - g10_hat - (g01_hat - g00_hat)) +
    (Tt * D / mean(D * Tt)) * (Y - g11_hat) -
    (D * (1 - Tt) / mean(D * (1 - Tt))) * (Y - g10_hat) -
    (m_hat * (1 - D) * Tt / (1 - m_hat)) /
    mean(m_hat * (1 - D) * Tt / (1 - m_hat)) * (Y - g01_hat) +
    (m_hat * (1 - D) * (1 - Tt) / (1 - m_hat)) /
    mean(m_hat * (1 - D) * (1 - Tt) / (1 - m_hat)) * (Y - g00_hat)
  
  theta_hat <- -mean(psi_b) / mean(psi_a)
  
  psi <- psi_a * theta_hat + psi_b
  influence_fun <- -1 / mean(psi_a) * psi
  
  if (is.null(cluster_var)) {
    se_hat <- sqrt(mean(psi^2) / (mean(psi_a)^2 * n))
    clustered <- FALSE
  } else {
    clusters <- unique(C)
    cluster_sums_squared <- rep(NA_real_, length(clusters))
    index <- 1
    for (cl in clusters) {
      cluster_sums_squared[index] <- sum(influence_fun[C == cl])^2
      index <- index + 1
    }
    var_hat <- 1 / n * sum(cluster_sums_squared)
    se_hat <- sqrt(var_hat / n)
    clustered <- TRUE
  }
  
  # ---- 4) Return object (base)
  out <- list(
    coefficients = c(ATT = theta_hat),
    se = se_hat,
    clustered = clustered,
    t = theta_hat / se_hat,
    p = 2 * (1 - pnorm(abs(theta_hat / se_hat))),
    ci_lower = theta_hat - 1.96 * se_hat,
    ci_upper = theta_hat + 1.96 * se_hat,
    N = n,
    N_treated = sum(D),
    N_control = sum(1 - D),
    N_post = sum(Tt),
    N_pre = sum(1 - Tt),
    y_var = y_var,
    d_var = d_var,
    t_var = t_var,
    x_vars = x_vars,
    n_covariates = ncol(X),
    cf = cf,
    methods = names(methods),
    feature_eng = feature_eng,
    cluster_var = cluster_var
  )
  
  # ------ 5) GATE  -------
  
  if (heterogeneity) {
    
    if (is.null(gate_var)) stop("heterogeneity = TRUE requires gate_var")
    
    G <- as.factor(data[[gate_var]])
    groups <- levels(G)
    
    # containers
    tau_hat <- rep(NA_real_, length(groups))
    names(tau_hat) <- groups
    
    IF_mat <- matrix(NA_real_, nrow = n, ncol = length(groups))
    colnames(IF_mat) <- groups
    
    for (j in seq_along(groups)) {
      g <- groups[j]
      I_g <- as.numeric(G == g)
      
      denom <- mean(I_g * psi_a)
      if (abs(denom) < 1e-12) {
        stop(paste0("Group ", g, " has near-zero denom E[I_g*psi_a]."))
      }
      
      num <- mean(I_g * psi_b)
      tau_hat[g] <- - num / denom
      
      # Influence function for tau_g
      psi_g <- I_g * (psi_a * tau_hat[g] + psi_b)
      IF_mat[, j] <- -1 / denom * psi_g
    }
    
    # SEs using the same logic as ATT, applied per group
    gate_se <- rep(NA_real_, length(groups))
    names(gate_se) <- groups
    
    if (is.null(cluster_var)) {
      for (j in seq_along(groups)) {
        gate_se[j] <- sqrt(mean(IF_mat[, j]^2) / n)
      }
      gate_clustered <- FALSE
    } else {
      clusters <- unique(C)
      for (j in seq_along(groups)) {
        infl <- IF_mat[, j]
        cluster_sums_squared <- rep(NA_real_, length(clusters))
        k <- 1
        for (cl in clusters) {
          cluster_sums_squared[k] <- sum(infl[C == cl])^2
          k <- k + 1
        }
        var_hat_g <- 1 / n * sum(cluster_sums_squared)
        gate_se[j] <- sqrt(var_hat_g / n)
      }
      gate_clustered <- TRUE
    }
    
    gate_t <- tau_hat / gate_se
    gate_p <- 2 * (1 - pnorm(abs(gate_t)))
    
    out$gate <- list(
      tau = tau_hat,
      se = gate_se,
      clustered = gate_clustered,
      t = gate_t,
      p = gate_p,
      ci_lower = tau_hat - 1.96 * gate_se,
      ci_upper = tau_hat + 1.96 * gate_se,
      groups = groups,
      N = n,
      gate_var = gate_var,
      influence_function = IF_mat
    )
  }
  class(out) <- "doubleML_did_cs_gate"
  return(out)
}


```


```{r}
# data preparation (was don alread in did_twfe, but I addedyq_numeric)
CashBail <- read_stata("CashBail.dta") %>%
  mutate(
    yq_numeric = as.numeric(as.factor(yq)),
    # Creating past_fel from prior_fel
    past_fel = as.numeric(prior_fel > 0),
    
    # Creating Denied
    Denied = as.numeric(initialbailtype == "Denied"),
    
    # Addressing NAs
    income_missing = as.numeric(is.na(defendantzipmedianhouseholdincom)),
    poverty_missing = as.numeric(is.na(PctBelPov)),
    
    defendantzipmedianhouseholdincom = ifelse(
      is.na(defendantzipmedianhouseholdincom),
      mean(defendantzipmedianhouseholdincom, na.rm = TRUE),
      defendantzipmedianhouseholdincom
    ),
    
    PctBelPov = ifelse(is.na(PctBelPov), mean(PctBelPov, na.rm = TRUE), PctBelPov),
    
    PostElig = EligibleOffense * Post,
    
    off = as.factor(off),
    yq = as.factor(yq),
    DOW = as.factor(DOW),
    CommissionerName_input = as.factor(CommissionerName_input),
    # add a group_variable that indicates whether a defendant is white and not hispanic,         black and not hispanic or hispanic
    ethnicity = factor(
      if_else(
        defendantisblack == 1 & Hisp == 0,
        "black_not_hisp",
        if_else(White == 1 & Hisp == 0, "white_not_hisp", 
                if_else(Hisp == 1, "hisp", NA))
      ),
      levels = c("black_not_hisp", "white_not_hisp", "hisp")
    )
  )
CashBail <- filter(CashBail, !is.na(ethnicity)) 


```

```{r}
# defining inputs
outcomes_table3 <- c("ROR", "Jail3days")
outcomes_table5 <- c("hasfta", "Rec6mo")
all_outcomes <- c(outcomes_table3, outcomes_table5)

covariates <- c(
  "defendantisblack",
  "Hisp",
  "White",
  "defendantageatarrest",
  "male",
  "defendantzipmedianhouseholdincom",
  "PctBelPov",
  "PD",
  "felony",
  "prior_FTA",
  "prior_9y",
  "past_fel",
  "income_missing",
  "poverty_missing"
)

# ML Methods
my_methods <- list(D.hat = list(
  m1 = list(method = "ranger_prop", tune = "fold"),
  m2 = list(method = "logit")
),

Y.hat.d = list(
  m1 = list(method = "ranger", tune = "fold"),
  m2 = list(method = "lasso")
))
```


```{r eval=FALSE}
# Store the estimation results for each outcome variable
dml_gate_results <- list()
set.seed(1234) 

# (ATT is automatically computed as part of GATT estimation)
for(y in all_outcomes) {
  message("Running ATT estimation for outcome: ", y)
  
  dml_results[[y]] <- doubleML_did_cs_gate(
    data = CashBail,
    y_var = y,
    d_var = "EligibleOffense",
    t_var = "Post",
    x_vars = covariates,
    methods = my_methods,
    cf = 5,              
    cluster_var = "off",
    heterogeneity = TRUE,
    gate_var = "ethnicity",
    feature_eng = TRUE
  )
}
```

The ROR coefficient estimate is 4 percentage points smaller than in the TWFE implementation, but still significant at the 1% level. Interestingly, in the TWFE implementation the coefficients for the jail, FTA and recidivism outcomes are all non-significant at the 10% level, while the Double ML results in all coefficients being significant at the 10% level, and the recidivism estimate even at the 5% level. 

The jail estimate of +3% has the opposite sign and is larger in absolute term than estimated through TWFE. Indeed a positive effect of the reform on the number of jailed individuals is surprising, as the reform would intuitively be expected to reduce the number of defendants jailed.  

The FTA and recidivism estimates both indicate a decrease of around 3%, which is stronger than estimated by TWFE. The authors propose that these positive effects could stem from the cash-bail having a very negative effect on financially vulnerable groups. For these groups having to pay hundreds or thousands of dollars in bail implies additional hardship and stress, which could decrease the likelihood of the defendant appearing at court and increase the chances of the defendants committing further crimes.  

Overall we can summarize that we found similar results through Double ML as also the authors found through TWFE. Only the coefficient for ROR is significantly different from zero, with the other 3 outcomes being very close to zero and harder to interpret.



#### DML DiD estimates of ATT and GATT

#### Table ?: ATT Results 
```{r eval=FALSE}

# Extract ATT from first GATT variable results (they all have the same ATT)
final_dml_table <- do.call(rbind, lapply(all_outcomes, function(y) {
  res <- dml_gate_results[[y]] # Take ATT from first gate_var
  data.frame(
    Outcome = y,
    ATT = res$coefficients["ATT"],
    Std_Error = res$se,
    t_stat = res$t,
    p_val = res$p,
    Lower_CI = res$ci_lower,
    Upper_CI = res$ci_upper
  )
}))

# make the  ATT table more beautiful
final_dm_rtable_beautiful <- final_dml_table %>%
  select(!c(Std_Error, t_stat)) %>%
  gt() %>%
  tab_header(
    title = "DiD ATT Results"
  ) %>%
  cols_label(
    Outcome = "Outcome",
    ATT = "ATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Outcome, ATT, p_val, Lower_CI, Upper_CI), decimals = 3) 

final_dm_rtable_beautiful
```

#### Heterogeneity
Furthermore we conducted a heterogeneity analysis.
Let $D_i \in \{0,1\}$ denote treatment status, $T_i \in \{0,1\}$ a time indicator, and $G_i \in \mathcal G$ a discrete group variable. We are interested in the group-specific average treatment effect on the treated $$
\tau_g \;=\; \mathbb E\!\left[ Y_i(1) - Y_i(0) \mid D_i=1,\, G_i=g \right].
$$ We start from a doubly robust score for the overall ATT, which yields $$
\theta^{ATT}
=
-\frac{\mathbb E[\psi_b]}{\mathbb E[\psi_a]}.
$$

To obtain a group-specific ATT, we do not condition on $D_i=1$. Instead, we define a group-weighted moment condition using the indicator $I_{g,i} = \mathbf 1\{G_i=g\}$: $$
\mathbb E\!\left[ I_{g,i}\,(\psi_{a,i}\,\tau_g + \psi_{b,i}) \right] = 0.
$$ Solving for $\tau_g$ gives $$
\boxed{
\tau_g
=
-\frac{\mathbb E[I_g\,\psi_b]}{\mathbb E[I_g\,\psi_a]}
}
$$ which is the GATT estimator implied by the same score that identifies the ATT.

#### GATT Results Summary
```{r eval=FALSE}
# Summary Table for GATT, dependent variable = hasfta
groups <- c("black_not_hisp", "white_not_hisp", "hisp")
summary_table_hasfta <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$hasfta$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$hasfta$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$hasfta$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$hasfta$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$hasfta$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$hasfta$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_hasfta) <- NULL


summary_table_hasfta_formatted <- summary_table_hasfta %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for hasfta"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 



```



```{r eval=FALSE}
# Summary Table for GATT, dependent variable = Rec6mo
summary_table_Rec6mo <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$Rec6mo$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_Rec6mo) <- NULL


summary_table_Rec6mo_formatted <- summary_table_Rec6mo %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for Rec6mo"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 

```


```{r eval=FALSE}
# Summary tabble DiD GATT, dep. variable = ROR
summary_table_ROR <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$ROR$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$ROR$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$ROR$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$ROR$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$ROR$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$ROR$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_ROR) <- NULL


summary_table_ROR_formatted <- summary_table_ROR %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for ROR"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 




```


```{r eval=FALSE}
# Summary Table for DiiD GATT, dep. variable = Jail3days
summary_table_Jail3days <- data.frame(
    Group = c("black", "white", "hispanic"),
    GATT = sapply(groups, function(x){dml_gate_results$Jail3days$gate$tau[[x]]}),
    Std_Error = sapply(groups, function(x){dml_gate_results$Jail3days$gate$se[[x]]}),
    t_stat = sapply(groups, function(x){dml_gate_results$Jail3days$gate$t[[x]]}),
    p_val = sapply(groups, function(x){dml_gate_results$Jail3days$gate$p[[x]]}),
    Lower_CI = sapply(groups, function(x){dml_gate_results$Jail3days$gate$ci_lower[[x]]}),
    Upper_CI = sapply(groups, function(x){dml_gate_results$Jail3days$gate$ci_upper[[x]]})
  ) %>% select(!c(t_stat, Std_Error))
rownames(summary_table_Jail3days) <- NULL


summary_table_Jail3days_formatted <- summary_table_Jail3days %>%
  gt() %>%
  tab_header(
    title = "DiD GATT Results for Jail3days"
  ) %>%
  cols_label(
    Group = "Group",
    GATT = "GATT",
    p_val= "P-Value",
    Lower_CI = "Lower_CI",
    Upper_CI = "Upper_CI"
  ) %>%
  fmt_number(columns = c(Group, GATT, p_val, Lower_CI, Upper_CI), decimals = 3) 



```

# Table ?: GATT for Dependent Variablehasfta
```{r eval=FALSE}
summary_table_hasfta_formatted 
```


# Table ?: GATT for Dependent Variable Rec6mo
```{r eval=FALSE}
summary_table_Rec6mo_formatted 
```


# Table ?: GATT for Dependent Variable ROR
```{r eval=FALSE}
summary_table_ROR_formatted 
```


# Table ?: GATT for Dependent Variable Jail
```{r eval=FALSE}
summary_table_Jail3days_formatted 
```
For brevity reasons we will here only interpret the heterogeneous effect estimates for the two violation outcomes, i.e. recidivism and failure to appear. Interestingly the estimates are similar for both of these outcomes. The reform seems to have been the most beneficial for white defendants, where the share of violators was lowered by 5 percentage points. Furthermore the share of black defendants violating the recidivism condition decreased by 2.6 percentage points. The other coefficients are far from statistically significant.


### 1.3.4 Reflection on DiD Assumptions

#### Policy Context

#### Event Study DiD

#### Endogenous Responses

#### Channels of Cash-Bail Reform Effects on Violations

## Figure 2: Jail Time Trend

```{r}
ggplot(subset(biweekly_avg, Outcome == "Jail"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "Jail",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly Jail Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

### Table x: Characterisitcs Composition

```{r}
covars_to_test <- c("defendantageatarrest", "male", "defendantisblack",
                    "prior_FTA", "prior_9y", "past_fel", "felony")
did_cov_data <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = as.numeric(prior_fel > 0),
    Released = if_else(Jail3days == 0, 1, 0)
  )

did_results <- map_df(covars_to_test, function(var) {
  form <- as.formula(paste(var, "~ Post * Released"))
  lm(form, data = did_cov_data) %>%
    tidy() %>%
    filter(term == "Post:Released") %>%
    mutate(Variable = var)
})

did_cov_table <- did_results %>%
  select(Variable, estimate, std.error, p.value) %>%
  mutate(
    Status = if_else(p.value > 0.05, "Stable", "Shift Detected")
  )


did_cov_table_formatted <- did_cov_table %>%
  gt() %>%
  tab_header(
    title = "DiD Test for Compositional Shifts"
  ) %>%
  cols_label(
    Variable = "Characteristic",
    estimate = "Interaction (Post × Released)",
    std.error = "Std. Error",
    p.value = "P-Value"
  ) %>%
  fmt_number(columns = c(estimate, std.error, p.value), decimals = 3) %>%
  tab_style(
    style = cell_text(color = "darkgreen", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "Stable")
  )

did_cov_table_formatted
```

### 1.3.5 Feedback
