---
title: 'Group Assignment: Replicating Ouss and Stevenson (2023)'
author: 'Group Members: Julian Schmidt, Laurens Mauz, Isa Afridi, Lukas Simon'
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: true
    toc_depth: 3
    number_sections: false
    theme: cosmo
    highlight: tango
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(fixest)
library(haven)
library(gt)
library(ggh4x)
library(modelsummary)
library(gtsummary)
library(DoubleML)
library(kableExtra)
```

```{r}
master_data <- read_stata("CashBail.dta")
master_data$baildate <- as.Date(master_data$baildate, origin = "1960-01-01")
```


## 1.1 Differences-in-Differences
## 1.2 Custom Double ML DiD Function
## 1.3 Replication of Ouss and Stevenson (2023)
### 1.3.1 Overview
### 1.3.2 Descriptives

#### Descriptive Statistics Table
```{r}
desc_stats_set <- master_data %>%
  filter(Post == 0) %>% 
  mutate(
    defendantzipmedianhouseholdincom = round(defendantzipmedianhouseholdincom, 0),
    PctBelPov = PctBelPov/100,
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = if_else(coalesce(prior_fel, 0) > 0, 1, 0),
    group_label = if_else(`EligibleOffense` == 1, "Eligible cases", "Ineligible cases")
  ) %>%
  select(
    group_label,
    defendantageatarrest, male, defendantisblack, Hisp, White, 
    defendantzipmedianhouseholdincom, PctBelPov, PD, felony, 
    prior_FTA, prior_9y, past_fel, 
    ROR, Nonmonetary, Unsecured, Bail_5000, Bail_HIGH, Denied, 
    Jail3days, hasfta, Rec6mo
  )

####
desc_stats_final <- desc_stats_set %>%
  tbl_summary(
    by = group_label,
    # forcing R to treat 0/1 variables as continuous so it can calculate mean 
    type = list(everything() ~ "continuous"), 
    label = list(
      defendantageatarrest ~ "Age",
      male ~ "Male",
      defendantisblack ~ "Black",
      Hisp ~ "Hispanic",
      White ~ "White",
      defendantzipmedianhouseholdincom ~ "Median household income in zip code",
      PctBelPov ~ "Fraction below poverty in zip code",
      PD ~ "Public Defender",
      felony ~ "Felony",
      prior_FTA ~ "Has a prior FTA",
      prior_9y ~ "Has a past conviction",
      past_fel ~ "Has a past felony conviction",
      ROR ~ "ROR",
      Nonmonetary ~ "Supervised release",
      Unsecured ~ "Unsecured monetary",
      Bail_5000 ~ "Secured bail up to 5,000",
      Bail_HIGH ~ "Secured bail over 5,000",
      Denied ~ "Denied bail",
      Jail3days ~ "Jail (3+ nights)",
      hasfta ~ "FTA",
      Rec6mo ~ "Recidivism"
    ),
    statistic = all_continuous() ~ "{mean}",
    digits = all_continuous() ~ 2,
    missing = "no"
  ) %>%
  modify_header(label = "**Variable**") %>%
  modify_spanning_header(all_stat_cols() ~ "**Pre-Reform Cases**") %>%
  bold_labels()

table1_final %>%
  as_gt() %>%
  tab_row_group(label = "Misconduct", rows = 20:21) %>%
  tab_row_group(label = "Pre-trial Conditions", rows = 13:19) %>%
  tab_row_group(label = "Defendant Characteristics", rows = 1:12)
```


#### Cash Bail System

```{r}
data_trends_ov <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),
    
    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

biweekly_avg <- data_trends_ov %>%
  group_by(W2, EligibleOffense) %>%
  summarise(
    ROR = mean(ROR, na.rm = TRUE),
    Jail = mean(Jail3days, na.rm = TRUE),
    FTA = mean(hasfta, na.rm = TRUE),
    Recidivism = mean(Rec6mo, na.rm = TRUE),
    .groups = "drop"
  ) %>%
 
  filter(W2 >= -12 & W2 <= 12 & W2 != 0) %>%
  pivot_longer(cols = c(ROR, Jail, FTA, Recidivism), names_to = "Outcome", values_to = "Rate")

scales_y <- list(
  Outcome == "ROR" ~ scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)),
  Outcome == "Jail" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "FTA" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)),
  Outcome == "Recidivism" ~ scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1))
)

biweekly_avg <- biweekly_avg %>%
  mutate(Outcome = factor(Outcome, levels = c("ROR", "Jail", "FTA", "Recidivism")))

# ggplot(biweekly_avg, aes(x = W2, y = Rate)) +
#   geom_point(color = "grey70", size = 1.5) +
#   
#   
#   geom_smooth(data = subset(biweekly_avg, W2 < 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_smooth(data = subset(biweekly_avg, W2 > 0), 
#               method = "lm", formula = y ~ poly(x, 2), 
#               se = FALSE, color = "black", size = 0.8) +
#   
#   geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
#   
#   facet_wrap(~Outcome, scales = "free_y") +
#   
#   facetted_pos_scales(y = scales_y) +
#   
#   scale_x_continuous(breaks = seq(-12, 12, 4), 
#                      labels = function(x) x / 2) +
#   
#   theme_minimal() +
#   labs(x = "Months relative to Feb. 22",
#        y = "Bi-weekly Rate") +
#   theme(panel.grid.minor = element_blank())
        
```


```{r}
# Figure 1a: ROR Time Trend
p_eligible <- ggplot(subset(biweekly_avg, Outcome == "ROR"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "ROR"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0.4, 0.8), breaks = seq(0.4, 0.8, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  
  theme_minimal() +
  labs(title = "ROR",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly ROR Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))
```

```{r}

# Figure 1c: ROR for Ineligible Cases
data_inel <- master_data %>%
  filter(EligibleOffense == 0) %>%
  mutate(
    days_rel = as.numeric(baildate - as.Date("2018-02-22")),

    W2 = ifelse(days_rel >= 0, floor(days_rel / 14), floor((days_rel + 1) / 14) - 1)
  )

# Bi-weekly ROR Rate
biweekly_ror_inel <- data_inel %>%
  group_by(W2) %>%
  summarise(Rate = mean(ROR, na.rm = TRUE), .groups = "drop") %>%
  filter(W2 >= -12 & W2 <= 11)

p_ineligible <- ggplot(biweekly_ror_inel, aes(x = W2, y = Rate)) +
 
  geom_point(color = "grey70", size = 1.2) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 < 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_smooth(data = filter(biweekly_ror_inel, W2 >= 0), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "grey30", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "ROR", 
       subtitle = "(Ineligible Cases)",
       x = "Months relative to Feb. 22", 
       y = "Bi-weekly rate") +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 12),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    panel.grid.minor = element_blank()
  )
  
```

## Figure 1a: ROR Time Trend & Figure 1b: ROR for Ineligible Cases
```{r}

combined_ror_plot <- p_eligible + p_ineligible


combined_ror_plot
```

#### Policy Reform
#### Target Parameters and Controls
### 1.3.3 Replication Results

#### DiD TWFE

```{r}
## Data Preparation
did_twfe <- master_data %>%
  mutate(
    # Creating past_fel from prior_fel 
    past_fel = as.numeric(prior_fel > 0),
    
    # Creating Denied 
    Denied = as.numeric(initialbailtype == "Denied"),
    
    # Addressing NAs
    income_missing = as.numeric(is.na(defendantzipmedianhouseholdincom)),
    poverty_missing = as.numeric(is.na(PctBelPov)),
    
    defendantzipmedianhouseholdincom = ifelse(is.na(defendantzipmedianhouseholdincom), 
                                              mean(defendantzipmedianhouseholdincom, na.rm = TRUE), 
                                              defendantzipmedianhouseholdincom),
    
    PctBelPov = ifelse(is.na(PctBelPov), 
                       mean(PctBelPov, na.rm = TRUE), 
                       PctBelPov),
    
    # Re-calculate or ensure PostElig is correct
    PostElig = EligibleOffense * Post,
  
    off = as.factor(off),
    yq = as.factor(yq),
    DOW = as.factor(DOW), 
    CommissionerName_input = as.factor(CommissionerName_input) 
  )

covariates <- c("defendantageatarrest", "male", "defendantisblack", "Hisp", 
                "White", "defendantzipmedianhouseholdincom", "PctBelPov", 
                "PD", "felony", "prior_FTA", "prior_9y", "past_fel", "income_missing", "poverty_missing")

# Column 1: ROR (No controls)
res_ror_1 <- feols(ROR ~ PostElig + EligibleOffense + Post, 
                   cluster = ~off, data = did_twfe)

# Column 2: ROR (With controls and Fixed Effects)
res_ror_2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, 
                   cluster = ~off, data = did_twfe)

# Column 3: Jail (No controls)
res_jail_3 <- feols(Jail3days ~ PostElig + EligibleOffense + Post, 
                    cluster = ~off, data = did_twfe)

# Column 4: Jail (With controls and Fixed Effects)
res_jail_4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, 
                    cluster = ~off, data = did_twfe)

## Pre-period Means & Models
mean_ror_pre <- CashBail %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(ROR, na.rm = TRUE)) %>% pull(mean)

mean_jail_pre <- CashBail %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Jail3days, na.rm = TRUE)) %>% pull(mean)

# Model (1): ROR No Controls
m1 <- feols(ROR ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): ROR With Controls & Fixed Effects
m2 <- feols(ROR ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

# Model (3): Jail No Controls
m3 <- feols(Jail3days ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Jail With Controls & Fixed Effects
m4 <- feols(Jail3days ~ PostElig + Post + .[covariates] | off + yq, cluster = ~off, data = did_twfe)

## SUMMARY TABLE


table_3 <- etable(m1, m2, m3, m4,
       headers = list("ROR" = 2, "Jail" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_ror_pre, 3), round(mean_ror_pre, 3), 
                             round(mean_jail_pre, 3), round(mean_jail_pre, 3))
       ),
      
       digits = 4, 
       digits.stats = "r0", 
       fitstat = ~n,     
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

mean_fta_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(hasfta, na.rm = TRUE)) %>% pull(mean)

mean_rec_pre <- master_data %>% 
  filter(EligibleOffense == 1, Post == 0) %>% 
  summarise(mean = mean(Rec6mo, na.rm = TRUE)) %>% pull(mean)

# Model (1): FTA No Controls
m5_1 <- feols(hasfta ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (2): FTA With Controls & Fixed Effects
m5_2 <- feols(hasfta ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

# Model (3): Recidivism No Controls
m5_3 <- feols(Rec6mo ~ PostElig + Post + EligibleOffense, cluster = ~off, data = did_twfe)

# Model (4): Recidivism With Controls & Fixed Effects
m5_4 <- feols(Rec6mo ~ PostElig + Post + .[covariates] | off + yq, 
              cluster = ~off, data = did_twfe)

#  SUMMARY TABLE 5
table_5 <- etable(m5_1, m5_2, m5_3, m5_4,
       headers = list("FTA" = 2, "Recidivism" = 2),
       dict = c(PostElig = "Eligible x Post 02/21"),
       keep = "Eligible x Post 02/21",
       extralines = list(
         "Controls"      = c("No", "Yes", "No", "Yes"),
         "Mean dep. var." = c(round(mean_fta_pre, 3), round(mean_fta_pre, 3), 
                             round(mean_rec_pre, 3), round(mean_rec_pre, 3))
       ),
       digits = 4, 
       digits.stats = "r0",
       fitstat = ~n,
       style.df = style.df(depvar.title = "", fixef.suffix = ""),
       notes = "Standard errors, clustered at the offense level, are in parentheses."
)

```

##  Table 3: DiD Estimates of policy effect on ROR and Jail & Table 5: DiD Estimates of policy effect on FTA and Recidivism 
```{r}
table_6 <- etable(
  m2, m4, m5_2, m5_4,
  dict = c(PostElig = "Eligible x Post 02/21"),
  keep = "Eligible x Post 02/21",
  extralines = list(
    "Mean dep. var." = c(round(mean_ror_pre, 3),
                         round(mean_jail_pre, 3),
                         round(mean_fta_pre, 3),
                         round(mean_rec_pre, 3))
  ),
  digits = 4,
  digits.stats = "r0",
  fitstat = ~n,
  style.df = style.df(depvar.title = "", fixef.suffix = ""),
  notes = "Standard errors, clustered at the offense level, are in parentheses."
)

# Render with kableExtra for a clean academic look
library(kableExtra)

table_6 %>%
  as.data.frame() %>%
  kbl(caption = "Table x: DiD Estimates of ROR, Jail, FTA, and Recidivism",
      booktabs = TRUE, 
      align = "lcccc",
      # This removes the extra column name row kable sometimes adds
      col.names = c("Characteristic", "ROR", "Jail", "FTA", "Recidivism")) %>%
  kable_classic(full_width = F, html_font = "Cambria") 

```


#### DiD Double ML
#### Heterogeneity
### 1.3.4 Reflection on DiD Assumptions
#### Policy Context
#### Event Study DiD
#### Endogenous Responses
#### Channels of Cash-Bail Reform Effects on Violations

## Figure 2: Jail Time Trend 
```{r}
ggplot(subset(biweekly_avg, Outcome == "Jail"), aes(x = W2, y = Rate)) +
  geom_point(color = "grey70", size = 1.5) +
  
  
  geom_smooth(data = subset(biweekly_avg, W2 < 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_smooth(data = subset(biweekly_avg, W2 > 0 & Outcome == "Jail"), 
              method = "lm", formula = y ~ poly(x, 2), 
              se = FALSE, color = "black", size = 0.8) +
  
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey30") +
  
  scale_y_continuous(limits = c(0, 0.4), breaks = seq(0, 0.4, 0.1)) +
  
  scale_x_continuous(breaks = seq(-12, 12, 4), 
                     labels = function(x) x / 2) +
  
  theme_minimal() +
  labs(title = "Jail",
    x = "Months relative to Feb. 22",
       y = "Bi-weekly Jail Rate") +
  theme(panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5))
```
### Table x: Characterisitcs Composition

```{r}
covars_to_test <- c("defendantageatarrest", "male", "defendantisblack", 
                    "prior_FTA", "prior_9y", "past_fel", "felony")
did_cov_data <- master_data %>%
  filter(EligibleOffense == 1) %>%
  mutate(
    Denied = if_else(initialbailtype == "Denied", 1, 0),
    past_fel = as.numeric(prior_fel > 0),
    Released = if_else(Jail3days == 0, 1, 0)
  )

did_results <- map_df(covars_to_test, function(var) {
  form <- as.formula(paste(var, "~ Post * Released"))
  lm(form, data = did_cov_data) %>%
    tidy() %>%
    filter(term == "Post:Released") %>% 
    mutate(Variable = var)
})

did_cov_table <- did_results %>%
  select(Variable, estimate, std.error, p.value) %>%
  mutate(
    Status = if_else(p.value > 0.05, "Stable", "Shift Detected")
  )


did_cov_table_formatted <- did_cov_table %>%
  gt() %>%
  tab_header(
    title = "DiD Test for Compositional Shifts"
  ) %>%
  cols_label(
    Variable = "Characteristic",
    estimate = "Interaction (Post × Released)",
    std.error = "Std. Error",
    p.value = "P-Value"
  ) %>%
  fmt_number(columns = c(estimate, std.error, p.value), decimals = 3) %>%
  tab_style(
    style = cell_text(color = "darkgreen", weight = "bold"),
    locations = cells_body(columns = Status, rows = Status == "Stable")
  )

did_cov_table_formatted
```


### 1.3.5 Feedback



